---
title: "Basic Recommender System"
author: "Cheryl Bowersox"
date: "Tuesday, February 14, 2017"
output: html_document
---


** Description of system:
This system provides recommendations of top 10 tasty beers for users based collaborative filtering.

Input: Matrix of users and rankings give to each beer across several attributes. 

Data Source: 

https://data.world/socialmediadata/beeradvocate

Unique Users are defined by variable review_profilename
Unique beer (items) are defined by variable beer_beerid
For this model I am only using the revew_overall rating to establish the recommendation, and only pulling the first 10K records. 

The function 'manualrec' was created not using any built in recommendation packages. This funciton takes the data table and a user name who the recommendation is for as input, and then creates a list of 10 recommended beers by finding the cosine similarity, pulling those identified users highest rated beers, and weighting the overall results by a factor of how many similar profiles listed that beer. In the case of a tie, the data is also sorted by most recent reviews. 
This manual function is inelegant, uses loops to create the final list, and is very resource-intensive to run. 


```{r message=FALSE, warning=FALSE}

# import libraries
library(recommenderlab)
library(dplyr)
library(tidyr)
library(reshape)

#read data
beerdata <- read.csv("~/GitHub/IS643/beerdata.csv")

#shape data to create matrix, profile X beer

beerrates <- beerdata %>%select(beer_beerid, review_profilename, review_overall)
df <- head(beerrates, 10000)%>%group_by(beer_beerid,review_profilename)%>%summarise(avgrev= mean(review_overall))

dfwide <- spread(df,beer_beerid, avgrev)

dfwide[is.na(dfwide)] <-0

# function to compute simliarity between vectors in this case comparing profile for user A with user 
#input X matrix the wide matrix, a username, output simliarity with other users

manualrec <- function(df, user){
  
  #calculate the affinity for this user's vector with all other users 
  #get users vector
  #user <- "abrand"
  u <- as.numeric(dfwide%>%filter(review_profilename == user)%>%select(-review_profilename))
  dfv <- dfwide%>%filter(review_profilename != user)
  dfrank <- data.frame(name = dfv$review_profilename, sim = -10)

  #awful loop use apply(?)
  for (i in 1:nrow(dfrank)){
    v <- as.numeric((dfv%>%select(-review_profilename))[i,])  
    dfrank$sim[i] <- crossprod(u, v)/(sqrt(crossprod(u)) * sqrt(crossprod(v))) 
  }
  # arrange users, top 5 users that match
  topten <- head(dfrank%>%arrange(desc(sim)),10)
  
  results <- beerdata%>%
    select(beer_name,review_profilename,review_overall,review_time)%>%
    filter(review_profilename %in% topten[,1])
  
  results <- merge(x = results, y = topten, by.x = "review_profilename",
                   by.y = "name", all.x = TRUE)
  #weight by count of recommendations
  results$beerrated <- results$review_overall * results$sim
  results <- results%>%select(beer_name,beerrated, review_time)%>%
    group_by(beer_name)%>%
    summarise(rank = mean(beerrated), count=n(), avgtime = mean(review_time))
  results$wrank <- results$rank * results$count
  
  #sort by max beerrated score, then by max review time
  recommend <- results%>%arrange(desc(wrank),desc(avgtime))%>%select(beer_name,wrank)
  
  return (head(recommend,10))
  }
  
```
For the first example I used user named "abrand" and found the related list

```{r}
(manualrec(dfwide,"abrand"))
```

A second recommendation example I used the user named "beerguy101"
```{r}
(manualrec(dfwide,"beerguy101"))
```

Using the built-in functions in the recommendlab package I get different, but much faster results:

```{r}
#convert to matrix
beerrates <- beerdata %>%select(user = review_profilename, beer = beer_name, rate = review_overall)
df <- head(beerrates, 10000)

ratemat <- as(df,"realRatingMatrix")

train <- ratemat[1:150]  # trainng data for model
rec =Recommender(train,method="UBCF", 
      param=list(normalize = "Z-score",method="Cosine",nn=5, minRating=1))

```

Predicting top 10 for "abrand" user again, this time using the UBCF method of the Recommender function in the package  yields a different, and quicker,  result
```{r}
pre_for_abrand <- predict(rec, ratemat["abrand",], n=10)

(as(pre_for_abrand, "list"))

````

Similarly, the prediction for the user "beerguy101" has different outcomes. 

```{r}
pre_for_beerguy101 <- predict(rec, ratemat["beerguy101",], n=10)
(as(pre_for_beerguy101, "list"))

````


